---
title: "Bayesian linear regression"
author: "Quan Gan"
date: "2/19/2022"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(rstan)
library(tidyverse)
library(knitr)
library(ggplot2)
options(mc.cores = parallel::detectCores())

```

## Load data in
There are three data sets with low, intermediate, and high errors.
```{r}
data_low <- read_csv("../Data/Dataset_S1_Mar8.csv")
data_intermediate <- read_csv("../Data/Dataset_S2_Mar8.csv")
data_high <- read_csv("../Data/Dataset_S3_Mar8.csv")
```

## Bayesian linear regression

Using the Bayesian linear regression without error
The model setup is is stan/Bayesian_linear_model.stan

```{R}
model_without_error <- function(data, iter, chains, seed){
  rstan_options(auto_write = TRUE)
  stan_date <- list(N = nrow(data), Temperature = data$Temperature, D47 = data$D47)
  model <- stan(file = "stan/Bayesian_linear_model.stan", data = stan_date, iter = iter, chains = chains, seed = seed)
  
  result <- rstan::extract(model)
  Intercept_names <- c("Median Intercept", "Intercept Std. Error", "2.5%", "92.5%")
  Intercept_values <- c(summary(result$alpha)[3], summary(model)$summary[1, 2], summary(result$alpha)[2],summary(result$alpha)[5])
  names(Intercept_values) <- Intercept_names
  
  Slope_names <- c("Median Slope", "Slope Std. Error", "2.5%", "92.5%")
  Slope_values <- c(summary(result$beta)[3], summary(model)$summary[2, 2], summary(result$beta)[2],summary(result$beta)[5])
  names(Slope_values) <- Slope_names
  
  return(c(Intercept_values, Slope_values))
}

# create data visualization
createPlot <- function(data, result, title, caption) {
  data %>% 
    ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
    geom_ribbon(aes(ymin = result[7] * x_TRUE + result[4] , ymax = result[8] * x_TRUE + result[3]), fill = "orange") +
    geom_point() +
    geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
    geom_abline(slope = result[5], intercept = result[1], color = 'yellow') + 
    labs(title = title,
         caption = caption) + 
    xlab("True Temperature") + 
    ylab("True D47")
}

```


### Data with low error
```{r}
results_low <- model_without_error(data_low, 1000, 4, 1234) 
rbind(results_low) %>% kable()
```



#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_low, 
           results_low, 
           "Bayesian model with no error in model and the low error data", 
           "Bayesian model with no error in model and the low error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```

### Data with intermediate error
```{r}
results_intermediate <- model_without_error(data_intermediate, 1000, 4, 1234) 
rbind(results_intermediate) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_intermediate, 
           results_intermediate, 
           "Bayesian model with no error in model and the intermediate error data", 
           "Bayesian model with no error in model and the intermediate error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```

### Data with high error
```{r}
results_high <- model_without_error(data_high, 1000, 4, 1234) 
rbind(results_high) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_high, 
           results_high, 
           "Bayesian model with no error in model and the high error data", 
           "Bayesian model with no error in model and the high error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```

## Bayesian linear regression Results summary
```{r}
rbind(results_low, results_intermediate, results_high) %>% kable()
```

## Bayesian Linear Regression with Errors

Using the Bayesian linear regression with error
The model setup is is stan/Bayesian_linear_model_with_error.stan

```{r}
model_with_error <- function(data, iter, chains, seed){
  rstan_options(auto_write = TRUE)
  stan_date <- list(N = nrow(data), 
                      Temperature = data$Temperature, 
                      Temperature_true = data$x_TRUE,
                      Temperature_error = abs(data$TempError),
                      D47 = data$D47,
                      D47_true = data$y_TRUE,
                      D47_error = abs(data$D47error))
  model <- stan(file = "stan/Bayesian_linear_model_with_error.stan", data = stan_date, iter = iter, chains = chains, seed = seed)
  
  result <- rstan::extract(model)
  Intercept_names <- c("Median Intercept", "Intercept Std. Error", "2.5%", "92.5%")
  Intercept_values <- c(summary(result$alpha)[3], summary(model)$summary[1, 2], summary(result$alpha)[2],summary(result$alpha)[5])
  names(Intercept_values) <- Intercept_names
  
  Slope_names <- c("Median Slope", "Slope Std. Error", "2.5%", "92.5%")
  Slope_values <- c(summary(result$beta)[3], summary(model)$summary[2, 2], summary(result$beta)[2],summary(result$beta)[5])
  names(Slope_values) <- Slope_names
  
  return(c(Intercept_values, Slope_values))
}
```

### Data with low error
```{r}
results_low <- model_with_error(data_low, 1000, 4, 1234) 
rbind(results_low) %>% kable()

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_low, 
           results_low, 
           "Bayesian model with error in model and the low error data", 
           "Bayesian model with error in model and the low error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```


### Data with Intermediate error
```{r}
results_intermediate <- model_with_error(data_intermediate, 1000, 4, 1234) 
rbind(results_intermediate) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_intermediate, 
           results_intermediate, 
           "Bayesian model with error in model and the intermediate error data", 
           "Bayesian model with error in model and the intermediate error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```

### Data with High error
```{r}
results_high <- model_with_error(data_high, 1000, 4, 1234) 
rbind(results_high) %>% kable()

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_high, 
           results_high, 
           "Bayesian model with error in model and the high error data", 
           "Bayesian model with error in model and the high error data. Red line as true coefficient. Yellow is Bayesian fitted coefficient")
```

## Bayesian linear regression Results summary
```{r}
rbind(results_low, results_intermediate, results_high) %>% kable()
```

