---
title: "Bayesian linear regression"
author: "Quan Gan"
date: "2/19/2022"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(rstan)
library(tidyverse)
library(knitr)
library(ggplot2)
options(mc.cores = parallel::detectCores())

```

## Load data in
There are three data sets with low, intermediate, and high errors.
```{r}
data_low <- read_csv("../Data/Dataset_S1_Mar8.csv")
data_intermediate <- read_csv("../Data/Dataset_S2_Mar8.csv")
data_high <- read_csv("../Data/Dataset_S3_Mar8.csv")
```

## Bayesian linear regression

Using the Bayesian linear regression without error
The model setup is is stan/Bayesian_linear_model.stan

### Data with low error
```{r}
stan_data_low <- list(N = nrow(data_low), Temperature = data_low$Temperature, D47 = data_low$D47)

rstan_options(auto_write = TRUE)
model_low <- stan(file = "stan/Bayesian_linear_model.stan", data = stan_data_low, iter = 1000, chains = 4, seed = 1234)

results_low <- rstan::extract(model_low)
cat("Median Intercept", median(results_low$alpha))
cat("Median Slope", median(results_low$beta))
summary(model_low)$summary
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_low %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_low)$summary[2,4] * x_TRUE + summary(model_low)$summary[1,8] , ymax = summary(model_low)$summary[2,8] * x_TRUE + summary(model_low)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(results_low$beta), intercept = median(results_low$alpha), color = 'yellow') + 
  labs(title = "Bayesian model with low error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### Data with intermediate error
```{r}
stan_data_intermediate <- list(N = nrow(data_intermediate), Temperature = data_intermediate$Temperature, D47 = data_intermediate$D47)

rstan_options(auto_write = TRUE)
model_intermediate <- stan(file = "stan/Bayesian_linear_model.stan", data = stan_data_intermediate, iter = 1000, chains = 4, seed = 1234)

results_intermediate <- rstan::extract(model_intermediate)
cat("Median Intercept", median(results_intermediate$alpha))
cat("Median Slope", median(results_intermediate$beta))
summary(model_intermediate)$summary
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_intermediate %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_intermediate)$summary[2,4] * x_TRUE + summary(model_intermediate)$summary[1,8] , ymax = summary(model_intermediate)$summary[2,8] * x_TRUE + summary(model_intermediate)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(results_intermediate$beta), intercept = median(results_intermediate$alpha), color = 'yellow') +
  labs(title = "Bayesian model with the intermediate error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### Data with high error
```{r}
stan_data_high <- list(N = nrow(data_high), Temperature = data_high$Temperature, D47 = data_high$D47)

rstan_options(auto_write = TRUE)
model_high <- stan(file = "stan/Bayesian_linear_model.stan", data = stan_data_high, iter = 1000, chains = 4, seed = 1234)

results_high <- rstan::extract(model_high)
cat("Median Intercept", median(results_high$alpha))
cat("Median Slope", median(results_high$beta))
summary(model_high)$summary
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_high %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_high)$summary[2,4] * x_TRUE + summary(model_high)$summary[1,8] , ymax = summary(model_high)$summary[2,8] * x_TRUE + summary(model_high)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(results_high$beta), intercept = median(results_high$alpha), color = 'yellow') +
  labs(title = "Bayesian model with the high error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

## Bayesian linear regression Results summary
```{r}
medianIntercept <- c(median(results_low$alpha), median(results_intermediate$alpha), median(results_high$alpha))
se_mean_Intercept <- c(summary(model_low)$summary[1,2], summary(model_intermediate)$summary[1,2], summary(model_high)$summary[1,2])
medianSlope <- c(median(results_low$beta), median(results_intermediate$beta), median(results_high$beta))
se_mean_Slope <- c(summary(model_low)$summary[2,2], summary(model_intermediate)$summary[2,2], summary(model_high)$summary[2,2])

mat <- array(c(medianIntercept, se_mean_Intercept, medianSlope, se_mean_Slope), dim = c(3, 4))
colnames(mat) <- c("median_Intercept", "se_mean_Intercept", "median_Slope", "se_mean_slope")
rownames(mat) <- c("Low Error", "Intermediate Error", "High Error")

mat %>% 
  kable()
```

## Bayesian Linear Regression with Errors

Using the Bayesian linear regression with error
The model setup is is stan/Bayesian_linear_model_with_error.stan

### Data with low error
```{r}
stan_data_low <- list(N = nrow(data_low), 
                      Temperature = data_low$Temperature, 
                      Temperature_true = data_low$x_TRUE,
                      Temperature_error = abs(data_low$TempError),
                      D47 = data_low$D47,
                      D47_true = data_low$y_TRUE,
                      D47_error = abs(data_low$D47error))


rstan_options(auto_write = TRUE)
model_low <- stan(file = "stan/Bayesian_linear_model_with_error.stan", data = stan_data_low, iter = 1000, chains = 4, seed = 1234)

cat("Median Intercept", summary(model_low)$summary[1,6])
cat("Median Slope", summary(model_low)$summary[2,6])
summary(model_low)$summary

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_low %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_low)$summary[2,4] * x_TRUE + summary(model_low)$summary[1,8] , ymax = summary(model_low)$summary[2,8] * x_TRUE + summary(model_low)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = summary(model_low)$summary[2,6], intercept = summary(model_low)$summary[1,6], color = 'yellow') + 
  labs(title = "Bayesian model with parameters' errors and low error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```


### Data with Intermediate error
```{r}
stan_data_intermediate <- list(N = nrow(data_intermediate), 
                      Temperature = data_intermediate$Temperature, 
                      Temperature_true = data_intermediate$x_TRUE,
                      Temperature_error = abs(data_intermediate$TempError),
                      D47 = data_intermediate$D47,
                      D47_true = data_intermediate$y_TRUE,
                      D47_error = abs(data_intermediate$D47error))


rstan_options(auto_write = TRUE)
model_intermediate <- stan(file = "stan/Bayesian_linear_model_with_error.stan", data = stan_data_intermediate, iter = 1000, chains = 4, seed = 1234)

cat("Median Intercept", summary(model_intermediate)$summary[1,6])
cat("Median Slope", summary(model_intermediate)$summary[2,6])
summary(model_intermediate)$summary

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_intermediate %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_intermediate)$summary[2,4] * x_TRUE + summary(model_intermediate)$summary[1,8] , ymax = summary(model_intermediate)$summary[2,8] * x_TRUE + summary(model_intermediate)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = summary(model_intermediate)$summary[2,6], intercept = summary(model_intermediate)$summary[1,6], color = 'yellow') + 
  labs(title = "Bayesian model with parameters' errors and intermediate error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```
### Data with High error
```{r}
stan_data_high <- list(N = nrow(data_high), 
                      Temperature = data_high$Temperature, 
                      Temperature_true = data_high$x_TRUE,
                      Temperature_error = abs(data_high$TempError),
                      D47 = data_high$D47,
                      D47_true = data_high$y_TRUE,
                      D47_error = abs(data_high$D47error))


rstan_options(auto_write = TRUE)
model_high <- stan(file = "stan/Bayesian_linear_model_with_error.stan", data = stan_data_high, iter = 1000, chains = 4, seed = 1234)

cat("Median Intercept", summary(model_high)$summary[1,6])
cat("Median Slope", summary(model_high)$summary[2,6])
summary(model_high)$summary

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_high %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = summary(model_high)$summary[2,4] * x_TRUE + summary(model_high)$summary[1,8] , ymax = summary(model_high)$summary[2,8] * x_TRUE + summary(model_high)$summary[1,4]), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = summary(model_high)$summary[2,6], intercept = summary(model_high)$summary[1,6], color = 'yellow') + 
  labs(title = "Bayesian model with parameters' errors and high error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

