---
title: "York and Diming Regression"
author: "Quan Gan"
date: "2/26/2022"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(knitr)
library(ggplot2)
library(IsoplotR)
library(boot)
library(deming)
```

## Load data in
There are three data sets with low, intermediate, and high errors.
```{r}
data_low <- read_csv("../Data/Dataset_S1.csv")
data_intermediate <- read_csv("../Data/Dataset_S2.csv")
data_high <- read_csv("../Data/Dataset_S3.csv")
```

## The functions for bootstrapping 1000 replicates with York models
```{r}
# https://www.statmethods.net/advstats/bootstrapping.html
# Boot library docs
# calculate the Intercept
York_model <- function(data, indices) {
  d <- data[indices,]                   # allows boot to select sample
  mark <- sample(                       # Mark for train-test
    c(rep(0, 0.7*nrow(d)), 
    c(rep(1, 0.3*nrow(d)))))      
  train <- d[mark == 0,]                # train data
  test <- d[mark == 1,]                 # test data
  york_data <- cbind(train$Temperature, train$TempError, train$D47, train$D47error)
  fit <- york(york_data)                # the York model
  Intercept <- fit$a['a']
  Slope <- fit$b['b']
  train_MSE <- mean((train$D47 - (train$Temperature * Slope + Intercept))^2)  # Train set MSE
  test_MSE <- mean((test$D47 - (test$Temperature * Slope + Intercept))^2) # Test set MSE
  return(c(Intercept, Slope, train_MSE, test_MSE))                    # get intercept
}

Deming_model <- function(formula, data, indices) {
  d <- data[indices,]                   # allows boot to select sample
  mark <- sample(                       # Mark for train-test
    c(rep(0, 0.7*nrow(d)), 
    c(rep(1, 0.3*nrow(d)))))      
  train <- d[mark == 0,]                # train data
  test <- d[mark == 1,]                 # test data
  fit <- deming(formula, data=train, xstd=abs(TempError), ystd=abs(D47error))    # the Deming model
  Intercept <- fit$coefficients[1]
  Slope <- fit$coefficients[2]
  train_MSE <- mean((train$D47 - (train$Temperature * Slope + Intercept))^2)  # Train set MSE
  test_MSE <- mean((test$D47 - (test$Temperature * Slope + Intercept))^2) # Test set MSE
  return(c(Intercept, Slope, train_MSE, test_MSE))                    # get intercept
}

```



## York Regresion

Unified regression algorithm. Given n pairs of measurements of x and y with their 
uncertainties x_error, y_error. The york function finds the best straight line 
using the least-squares algorithm of York.

### Low error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
york_results_low <- boot(data=data_low, 
                statistic=York_model,
                R=1000)


cat("Median Intercept:", median(york_results_low$t[,1]))
cat("Intercept Std. Error:", sd(york_results_low$t[,1]))
cat("Meidan Slope:", median(york_results_low$t[,2]))
cat("Slope Std. Error:", sd(york_results_low$t[,2]))
cat("Median Train MSE:", median(york_results_low$t[,3]))
cat("Meidan Test MSE:", median(york_results_low$t[,4]))
print(boot.ci(york_results_low, type="bca", index=1))
print(boot.ci(york_results_low, type="bca", index=2))
```
#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_low %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0340 * x_TRUE + 0.3004  , ymax = 0.0433  * x_TRUE + 0.1925), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(york_results_low$t[,2]), intercept = median(york_results_low$t[,1]), color = 'yellow') + 
  labs(title = "York model with the low error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### Intermediate error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
york_results_intermediate <- boot(data=data_intermediate, 
                statistic=York_model,
                R=1000)


cat("Median Intercept:", median(york_results_intermediate$t[,1]))
cat("Intercept Std. Error:", sd(york_results_intermediate$t[,1]))
cat("Meidan Slope:", median(york_results_intermediate$t[,2]))
cat("Slope Std. Error:", sd(york_results_intermediate$t[,2]))
cat("Median Train MSE:", median(york_results_intermediate$t[,3]))
cat("Meidan Test MSE:", median(york_results_intermediate$t[,4]))
print(boot.ci(york_results_intermediate, type="bca", index=1))
print(boot.ci(york_results_intermediate, type="bca", index=2))
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_intermediate %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0316 * x_TRUE + 0.3321   , ymax = 0.0455   * x_TRUE + 0.1683), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(york_results_intermediate$t[,2]), intercept = median(york_results_intermediate$t[,1]), color = 'yellow') + 
  labs(title = "York model with the intermediate error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### High error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
york_results_high <- boot(data=data_high, 
                statistic=York_model,
                R=1000)


cat("Median Intercept:", median(york_results_high$t[,1]))
cat("Intercept Std. Error:", sd(york_results_high$t[,1]))
cat("Meidan Slope:", median(york_results_high$t[,2]))
cat("Slope Std. Error:", sd(york_results_high$t[,2]))
cat("Median Train MSE:", median(york_results_high$t[,3]))
cat("Meidan Test MSE:", median(york_results_high$t[,4]))
print(boot.ci(york_results_high, type="bca", index=1))
print(boot.ci(york_results_high, type="bca", index=2))
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_high %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0435 * x_TRUE + 0.1889    , ymax = 0.0644  * x_TRUE -0.0610), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(york_results_high$t[,2]), intercept = median(york_results_high$t[,1]), color = 'yellow') + 
  labs(title = "York model with the high error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

## Median slopes/intercepts VS true parameters

Mean slope/Intercept VS Median slope/Intercept
```{r}
medianIntercept <- c(median(york_results_low$t[,1]), median(york_results_intermediate$t[,1]), median(york_results_high$t[,1]))
sdIntercept <- c(sd(york_results_low$t[,1]), sd(york_results_intermediate$t[,1]), sd(york_results_high$t[,1]))
medianSlope <- c(median(york_results_low$t[,2]), median(york_results_intermediate$t[,2]), median(york_results_high$t[,2]))
sdSlope <- c(sd(york_results_low$t[,2]), sd(york_results_intermediate$t[,2]), sd(york_results_high$t[,2]))
medianTrainMSE <- c(median(york_results_low$t[,3]), median(york_results_intermediate$t[,3]), median(york_results_high$t[,3]))
medianTestMSE<- c(median(york_results_low$t[,4]), median(york_results_intermediate$t[,4]), median(york_results_high$t[,4]))

mat <- array(c(medianIntercept, sdIntercept, medianSlope, sdSlope, medianTrainMSE, medianTestMSE), dim = c(3, 6))
colnames(mat) <- c("median_Intercept", "sd_Intercept", "median_Slope", "sd_slope", "median_train_MSE", "median_test_MSE")
rownames(mat) <- c("Low Error", "Intermediate Error", "High Error")
mat %>% 
  kable()
```

## Deming Regression

Find the MLE line relating x and y when both are measured with error. When the variances of x and y are constant and equal, this is the special case of Deming regression.

### Low error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
deming_results_low <- boot(data=data_low, 
                statistic=Deming_model,
                R=1000,
                formula=D47~Temperature)


cat("Median Intercept:", median(deming_results_low$t[,1]))
cat("Intercept Std. Error:", sd(deming_results_low$t[,1]))
cat("Meidan Slope:", median(deming_results_low$t[,2]))
cat("Slope Std. Error:", sd(deming_results_low$t[,2]))
cat("Median Train MSE:", median(deming_results_low$t[,3]))
cat("Meidan Test MSE:", median(deming_results_low$t[,4]))
print(boot.ci(deming_results_low, type="bca", index=1))
print(boot.ci(deming_results_low, type="bca", index=2))
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_low %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0340 * x_TRUE + 0.3005  , ymax = 0.0435  * x_TRUE + 0.1897), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(deming_results_low$t[,2]), intercept = median(deming_results_low$t[,1]), color = 'yellow') + 
  labs(title = "Deming model with the low error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### Intermediate error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
deming_results_intermediate <- boot(data=data_intermediate, 
                statistic=Deming_model,
                R=1000,
                formula=D47~Temperature)


cat("Median Intercept:", median(deming_results_intermediate$t[,1]))
cat("Intercept Std. Error:", sd(deming_results_intermediate$t[,1]))
cat("Meidan Slope:", median(deming_results_intermediate$t[,2]))
cat("Slope Std. Error:", sd(deming_results_intermediate$t[,2]))
cat("Median Train MSE:", median(deming_results_intermediate$t[,3]))
cat("Meidan Test MSE:", median(deming_results_intermediate$t[,4]))
print(boot.ci(deming_results_intermediate, type="bca", index=1))
print(boot.ci(deming_results_intermediate, type="bca", index=2))
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_intermediate %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0314 * x_TRUE + 0.3397   , ymax = 0.0455  * x_TRUE + 0.1684), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(deming_results_intermediate$t[,2]), intercept = median(deming_results_intermediate$t[,1]), color = 'yellow') + 
  labs(title = "Deming model with the intermediate error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

### High error data
```{r}
# set seed for production
set.seed(1234)

# bootstrapping with 1000 replications
deming_results_high <- boot(data=data_high, 
                statistic=Deming_model,
                R=1000,
                formula=D47~Temperature)


cat("Median Intercept:", median(deming_results_high$t[,1]))
cat("Intercept Std. Error:", sd(deming_results_high$t[,1]))
cat("Meidan Slope:", median(deming_results_high$t[,2]))
cat("Slope Std. Error:", sd(deming_results_high$t[,2]))
cat("Median Train MSE:", median(deming_results_high$t[,3]))
cat("Meidan Test MSE:", median(deming_results_high$t[,4]))
print(boot.ci(deming_results_high, type="bca", index=1))
print(boot.ci(deming_results_high, type="bca", index=2))
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
data_high %>% 
  ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
  geom_ribbon(aes(ymin = 0.0435 * x_TRUE + 0.1889 , ymax = 0.0644   * x_TRUE - 0.0609), fill = "orange") +
  geom_point() +
  geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
  geom_abline(slope = median(deming_results_high$t[,2]), intercept = median(deming_results_high$t[,1]), color = 'yellow') + 
  labs(title = "Deming model with the high error data") + 
  xlab("True Temperature") + 
  ylab("True D47")
```

## Median slopes/intercepts VS true parameters

Mean slope/Intercept VS Median slope/Intercept
```{r}
medianIntercept <- c(median(deming_results_low$t[,1]), median(deming_results_intermediate$t[,1]), median(deming_results_high$t[,1]))
sdIntercept <- c(sd(deming_results_low$t[,1]), sd(deming_results_intermediate$t[,1]), sd(deming_results_high$t[,1]))
medianSlope <- c(median(deming_results_low$t[,2]), median(deming_results_intermediate$t[,2]), median(deming_results_high$t[,2]))
sdSlope <- c(sd(deming_results_low$t[,2]), sd(deming_results_intermediate$t[,2]), sd(deming_results_high$t[,2]))
medianTrainMSE <- c(median(deming_results_low$t[,3]), median(deming_results_intermediate$t[,3]), median(deming_results_high$t[,3]))
medianTestMSE<- c(median(deming_results_low$t[,4]), median(deming_results_intermediate$t[,4]), median(deming_results_high$t[,4]))

mat <- array(c(medianIntercept, sdIntercept, medianSlope, sdSlope, medianTrainMSE, medianTestMSE), dim = c(3, 6))
colnames(mat) <- c("median_Intercept", "sd_Intercept", "median_Slope", "sd_slope", "median_train_MSE", "median_test_MSE")
rownames(mat) <- c("Low Error", "Intermediate Error", "High Error")
mat %>% 
  kable()
```