---
title: "York and Diming Regression"
author: "Quan Gan"
date: "2/26/2022"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(knitr)
library(ggplot2)
library(IsoplotR)
library(boot)
library(deming)
```

## Load data in
There are three data sets with low, intermediate, and high errors.
```{r}
data_low <- read_csv("../Data/Dataset_S1_Mar8.csv")
data_intermediate <- read_csv("../Data/Dataset_S2_Mar8.csv")
data_high <- read_csv("../Data/Dataset_S3_Mar8.csv")
```

## The functions for bootstrapping 1000 replicates with York models
```{r}
# https://www.statmethods.net/advstats/bootstrapping.html
# Boot library docs
# calculate the Intercept
York_model <- function(formula, data, indices) {
  d <- data[indices,]                   # allows boot to select sample
  mark <- sample(                       # Mark for train-test
    c(rep(0, 0.7*nrow(d)), 
    c(rep(1, 0.3*nrow(d)))))      
  train <- d[mark == 0,]                # train data
  test <- d[mark == 1,]                 # test data
  york_data <- cbind(train$Temperature, train$TempError, train$D47, train$D47error)
  fit <- york(york_data)                # the York model
  Intercept <- fit$a['a']
  Slope <- fit$b['b']
  train_MSE <- mean((train$D47 - (train$Temperature * Slope + Intercept))^2)  # Train set MSE
  test_MSE <- mean((test$D47 - (test$Temperature * Slope + Intercept))^2) # Test set MSE
  return(c(Intercept, Slope, train_MSE, test_MSE))                    # get intercept
}

Deming_model <- function(formula, data, indices) {
  d <- data[indices,]                   # allows boot to select sample
  mark <- sample(                       # Mark for train-test
    c(rep(0, 0.7*nrow(d)), 
    c(rep(1, 0.3*nrow(d)))))      
  train <- d[mark == 0,]                # train data
  test <- d[mark == 1,]                 # test data
  fit <- deming(formula, data=train, xstd=abs(TempError), ystd=abs(D47error))    # the Deming model
  Intercept <- fit$coefficients[1]
  Slope <- fit$coefficients[2]
  train_MSE <- mean((train$D47 - (train$Temperature * Slope + Intercept))^2)  # Train set MSE
  test_MSE <- mean((test$D47 - (test$Temperature * Slope + Intercept))^2) # Test set MSE
  return(c(Intercept, Slope, train_MSE, test_MSE))                    # get intercept
}

model <- function(seed, statistic, dataSet, formula, replicate){
  set.seed(seed)
  results <- boot(data = dataSet,
                  statistic = statistic,
                  R = replicate,
                  formula = formula)
  
  CI_Intecept <- boot.ci(results, type="bca", index=1)
  Intercept_names <- c("Median Intercept", "Intercept Std. Error", "2.5%", "92.5%")
  Intercept_values <- c(median(results$t[,1]), sd(results$t[,1]) / sqrt(length(results$t[,1])), CI_Intecept$bca[4], CI_Intecept$bca[5])
  names(Intercept_values) <- Intercept_names
  
  CI_Slope <- boot.ci(results, type="bca", index=2)
  Slope_names <- c("Median Slope", "Slope Std. Error", "2.5%", "92.5%")
  Slope_values <- c(median(results$t[,2]), sd(results$t[,2]) / sqrt(length(results$t[,2])), CI_Slope$bca[4], CI_Slope$bca[5])
  names(Slope_values) <- Slope_names
  
  MSE_names <- c("Median Train MSE", "Median Test MSE")
  MSE_values <- c(median(results$t[,3]), median(results$t[,4]))
  names(MSE_values) <- MSE_names
  
  
  Intercept_values %>% kable()
  Slope_values %>% kable()
  MSE_values %>% kable()
  
  return (c(Intercept_values, Slope_values, MSE_values))
}

# create data visualization
createPlot <- function(data, result, title, caption) {
  data %>% 
    ggplot(aes(x = x_TRUE, y=y_TRUE)) + 
    geom_ribbon(aes(ymin = result[7] * x_TRUE + result[4] , ymax = result[8] * x_TRUE + result[3]), fill = "orange") +
    geom_point() +
    geom_abline(slope = 0.0369, intercept = 0.268, color = 'red') +
    geom_abline(slope = result[5], intercept = result[1], color = 'yellow') + 
    labs(title = title,
         caption = caption) + 
    xlab("True Temperature") + 
    ylab("True D47")
}

```



## York Regresion

Unified regression algorithm. Given n pairs of measurements of x and y with their 
uncertainties x_error, y_error. The york function finds the best straight line 
using the least-squares algorithm of York.

### Low error data
```{r}
york_result_low <- model(1234, York_model, data_low, D47~Temperature, 1000)

rbind(york_result_low) %>% kable()
```
#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_low, 
           york_result_low, 
           "York model with the low error data", 
           "York model with the low error data. Red line as true coefficient. Yellow is York fitted coefficient")
```

### Intermediate error data
```{r}
york_result_intermediate <- model(1234, York_model, data_intermediate, D47~Temperature, 1000)

rbind(york_result_intermediate) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_intermediate, 
           york_result_intermediate, 
           "York model with the intermediate error data", 
           "York model with the intermediate error data. Red line as true coefficient. Yellow is York fitted coefficient")
```

### High error data
```{r}
york_result_high <- model(1234, York_model, data_high, D47~Temperature, 1000)

rbind(york_result_high) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_high, 
           york_result_high, 
           "York model with the high error data", 
           "York model with the high error data. Red line as true coefficient. Yellow is York fitted coefficient")
```

## Median slopes/intercepts VS true parameters

Mean slope/Intercept VS Median slope/Intercept
```{r}
rbind(york_result_low, york_result_intermediate, york_result_high) %>% kable()
```

## Deming Regression

Find the MLE line relating x and y when both are measured with error. When the variances of x and y are constant and equal, this is the special case of Deming regression.

### Low error data
```{r}
deming_result_low <- model(1234, Deming_model, data_low, D47~Temperature, 1000)

rbind(deming_result_low) %>% kable()

```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_low, 
           deming_result_low, 
           "Deming model with the low error data", 
           "Deming model with the low error data. Red line as true coefficient. Yellow is deming fitted coefficient")
```

### Intermediate error data
```{r}
deming_result_intermediate <- model(1234, Deming_model, data_intermediate, D47~Temperature, 1000)

rbind(deming_result_intermediate) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_intermediate, 
           deming_result_intermediate, 
           "Deming model with the intermediate error data", 
           "Deming model with the intermediate error data. Red line as true coefficient. Yellow is deming fitted coefficient")
```

### High error data
```{r}
deming_result_high <- model(1234, Deming_model, data_high, D47~Temperature, 1000)

rbind(deming_result_high) %>% kable()
```

#### Comparing result to the true parameters
Plot confident interval area of the median slope/intercept of bootstrapping results
And comparing it to the true slope/intercept

I use the true data points to plot.
```{r}
createPlot(data_high, 
           deming_result_high, 
           "Deming model with the high error data", 
           "Deming model with the high error data. Red line as true coefficient. Yellow is deming fitted coefficient")
```

## Median slopes/intercepts VS true parameters

Mean slope/Intercept VS Median slope/Intercept
```{r}
rbind(deming_result_low, deming_result_intermediate, deming_result_high) %>% kable()
```